Matching the Time of Emergence of Transboundary Fish Stocks to Lead Time for

### 2020-07-26 Scripts/Initial_Analysis.Rmd
Changed code so that the spp's centroid is the average of the girds >0.95 percentile

#### READ ME !!! ####
# Run this chunk before knit so you make sure you have all pkgs installed in R

@@

GetCentroids <- function(taxon_key,
                         year, 
                         data_type = "Catch",
                         ensemble, 
                         path,
                         Neighbours,
                         eez_centroid
                         ){

@@

  Spp_Dist <- read_dbem_ens(
    taxon_key = taxon_key,
    year = year,
    data_type = data_type,
    ensemble = ensemble
  )

@@

      # Centroid is defined as the tile with the higest species value
      Centroids <- Sub_Trans_Spp %>% 
        left_join(EEZ_ids, by = "eez_name") %>% 
        group_by(taxon_key,year,ensemble,neighbour_id) %>%  # for whole distribution
        top_n(1,value) %>% 
        distinct(value,year, .keep_all = TRUE) %>%  # removes duplicates when tile is in between EEZs
        left_join(Sub_EEZ_centroid, by = "taxon_key") %>%
        select(eez_name =eez_name.y, everything() ,-eez_name.x) %>% 
        # Estimate distances between spp centroid and EEZ centroid
        mutate(
          distance = geodist(longitude, # eez longitude
                             latitude,  # eez latitude
                             lon,  # species longitude
                             lat, # species latitude
                             units="km")
        ) %>% 
        select(taxon_key,eez_name,neighbour_id,ensemble,year,distance)

@@

    # Determine the ids
    Ids <- unique(Neighbours_combo$neighbour_id)

@@

  # Estimate noice from ensemble members
  historic_sd <- Trans_index %>% 
    filter(year < Hist_y) %>% 
    group_by(taxon_key) %>% 
    summarise_at(vars("RMean"),
                 c(hist_mean=mean,
                   noise=sd,
                   n_ens = ~n())
    )

@@

# Load all sp in one table
Eme_Res <- bind_rows(lapply(Paths, fread)) %>% 
  filter(
    n_ens == 10
  )
# warnings() for empty data species

@@

@@

# Per Stock
stock_d_hist <- Eme_Res %>% 
  group_by(emerging_yr,taxon_key,Tresh) %>%
  # summarise(n_tax = length(unique(eez_name))) %>%
  summarise(n()) %>% 
  mutate(
    level = "stock"
  ) 
  # ggplot() +
  # geom_histogram(
  #   aes(
  #     x = emerging_yr
  #   ),
  #   stat = "count"
  # )

# Per EEZs
eez_d_hist <- Eme_Res %>% 
  group_by(eez_name,emerging_yr,Tresh) %>% 
  # summarise(n_tax = length(unique(taxon_key))) %>% 
  summarise(n()) %>% 
  mutate(
    level = "EEZ"
  ) %>% 
  ungroup() %>% 
  select(1,3,4)

hist_data <- stock_d_hist %>% 
  bind_rows(eez_d_hist) %>% 
  filter(Tresh != "NA")

ggplot() +
  geom_density(data = subset(hist_data, level == "EEZ"),
    aes(
      x = emerging_yr,
      fill = level
    ),
    stat = "count",
    alpha= 0.8,
    size=0.5,
  ) +
  geom_density(data = subset(hist_data, level == "stock"),
    aes(
      x = emerging_yr,
      fill = level
    ),
    stat = "count",
    alpha= 0.8,
    size=0.5,
  ) +
    scale_y_continuous("Frequency",
                       breaks = seq(0,90,10),
                       limits = c(0,100)
    ) +
    scale_x_continuous("Year of emergence",
                       breaks = seq(2000,2100,10)
    ) +
    my_ggtheme_p() +
    scale_fill_viridis(discrete = T) +
  facet_wrap(~Tresh)

@@

##### Time Lapse plot

@@

# Per Stock
stock_d <-  Eme_Res %>% 
  left_join(Neighbours) %>% 
    distinct(taxon_key,neighbour_id, .keep_all = TRUE) %>% 
    group_by(taxon_key,Tresh) %>% 
    summarise(
      taxon_emer = min(emerging_yr)
    ) %>% 
    group_by(taxon_emer,Tresh) %>% 
  summarise(
    n_tax = n()
  ) %>% 
  mutate(
    com_sum = cumsum(n_tax),
    level = "Stock"
  ) %>% 
    select(1,3,4)

  #  EEZ data
eez_d <-  Eme_Res %>% 
  left_join(Neighbours) %>% 
  filter(n_ens == 10) %>% 
    distinct(eez_name,neighbour_id, .keep_all = TRUE) %>% 
    group_by(eez_name,Tresh) %>% 
    summarise(
      taxon_emer = min(emerging_yr)
    ) %>% 
    group_by(taxon_emer,Tresh) %>% 
  summarise(
    n_tax = n()
  ) %>% 
  mutate(
    com_sum = cumsum(n_tax),
    level = "EEZ"
  ) %>% 
    select(1,3,4)

# Plot Data
plot_data <- stock_d %>% 
  bind_rows(eez_d) %>% 
  filter(Tresh != "NA")

  # filter(emerging_yr >= 2020) %>% 
  ggplot() +
    geom_area(data = subset(plot_data, level == "Stock"),
              aes(
                x = as.numeric(taxon_emer),
                y = com_sum,
                fill = level
              ),
              alpha= 0.8,
              size=0.5,
              colour="black"
    )  +
    geom_area(data = subset(plot_data, level == "EEZ"),
              aes(
                x = taxon_emer,
                y = com_sum,
                fill = level
              ),
              alpha= 0.8,
              size=0.5,
              colour="black"
    ) +
    # Comment out for option A (Change plot name!!)
    geom_hline(yintercept = length(unique(Eme_Res$eez_name)), color = "grey",linetype = "dashed") +
  geom_text(aes(label="Total number of EEZs", x = 2010, y = length(unique(Eme_Res$eez_name))+10), color = "grey70",size =3) +
    geom_hline(yintercept = length(unique(Eme_Res$taxon_key)), color = "grey",linetype = "dashed") +
  geom_text(aes(label="Total number of Taxa", x = 2010, y = length(unique(Eme_Res$taxon_key))+10), color = "grey70",size =3) +
    scale_y_continuous("Cumulatuve Curve",
                       breaks = seq(0,500,50),
                       limits = c(0,500)
    ) +
    scale_x_continuous("Year of emergence",
                       breaks = seq(2000,2100,10)
    ) +
    my_ggtheme_p() +
    scale_fill_viridis(discrete = T) +
    facet_wrap(~Tresh)
  # ggsave("./Figures/figure_accumulative_curve_B.png",
  #        width = 6,
  #        height = 6,
  #        units = "in"
  # )
```

@@

>>>>

@@

GetCentroids <- function(taxon_key,
                         year, 
                         data_type = "Catch",
                         ensemble, 
                         path,
                         Neighbours,
                         eez_centroid,
                         p = 0.95 # percentile of data for central distribution
                         ){

@@

  Spp_Dist <- read_dbem_ens(
    taxon_key = taxon_key,
    year = year,
    data_type = data_type,
    ensemble = ensemble,
    path = my_path("G","DBEM")
  )

@@

      # Function to get percentiles
      # https://www.tidyverse.org/blog/2020/03/dplyr-1-0-0-summarise/
      quibble <- function(x, q = p) {
        tibble(x = quantile(x, q), q = q)
      }

      # Get the percentile treshold per year, ensemble and neighbour id
      percentile <- Sub_Trans_Spp %>% 
        left_join(EEZ_ids, by = "eez_name") %>% 
        group_by(taxon_key,year,ensemble,neighbour_id) %>%  # for whole distribution
        do(quibble(.$value, p))

       Centroids <- Sub_Trans_Spp %>% 
         left_join(EEZ_ids, by = "eez_name") %>% 
         left_join(percentile,
                   by = c("taxon_key","year","ensemble","neighbour_id")
                   ) %>% 
         mutate(to_filter = ifelse(value < x,"out","keep")) %>% 
         filter(to_filter == "keep") %>% 
         group_by(taxon_key,year,ensemble,neighbour_id) %>%  # for whole distribution
         summarise_at(vars(lon,lat),mean) %>% 
         left_join(Sub_EEZ_centroid, by = "taxon_key") %>%
         # select(eez_name =eez_name.y, everything() ,-eez_name.x) %>% 
         # Estimate distances between spp centroid and EEZ centroid
         mutate(
           distance = geodist(longitude, # eez longitude
                              latitude,  # eez latitude
                              lon,  # species longitude
                              lat, # species latitude
                              units="km")
         ) %>% 
         select(taxon_key,eez_name,neighbour_id,ensemble,year,distance)

@@

    }
    # Determine the ids
    Ids <- unique(Neighbours_combo$neighbour_id)

@@

  # Estimate noice from ensemble members
  historic_sd <- Trans_index %>% 
    filter(year < Hist_y) %>% 
    group_by(taxon_key,neighbour_id,ensemble) %>% #Average historic data per ensemble
    summarise_at(vars("trend"),
                 c(trend= mean)
                 ) %>% 
    group_by(taxon_key,neighbour_id) %>% #Capture ensemble variation
    summarise_at(vars("trend"),
                 c(hist_mean=mean,
                   noise=sd,
                   n_ens = ~n())
                 )

@@

# Load all sp in one table
Eme_Res <- bind_rows(lapply(Paths, fread)) %>% 
  filter(
    n_ens == 10
  ) %>% 
  left_join(Neighbours,
            by = c("eez_name","eez_neighbour"))
# warnings() for empty data species

@@

# Load it!
World_eez_sf <- st_read(dsn = path_world,
                    layer =file_path_sans_ext(fnam_world)) %>% 
  rename(eez_name = Name) %>% 
  st_transform(crs = 4326) %>% # 4326
  st_simplify(preserveTopology = TRUE, dTolerance = 100)

@@

# First time a Species becomes emergent
Eme_Res %>% 
  filter(
    n_ens == 10#,
    # Tresh == "Tresh_one"
  ) %>% 
  group_by(Tresh,taxon_key,emerging_yr) %>% 
  summarise(
    n_trans = length(unique(eez_name))
  ) %>% 
  group_by(Tresh,taxon_key) %>% 
  summarise(
    ntrans = min(emerging_yr)
  ) %>% 
  group_by(Tresh,ntrans) %>% 
  summarise(
    n_taxa = length(unique(taxon_key))
  ) %>% 
  arrange(-n_taxa) %>% 
  # View()
  # group_by(Tresh) %>% 
  # summarise(sum(n_taxa))
  ggplot() +
  geom_line(
    aes(
      x = ntrans,
      y = n_taxa,
      color = Tresh
    )
  )

# EEZ, first time an EEZ has a emergent stock
Eme_Res %>% 
  filter(
    n_ens == 10#,
    # Tresh == "Tresh_one"
  ) %>% 
  group_by(Tresh,eez_name,emerging_yr) %>% 
  summarise(
    n_trans = length(unique(taxon_key))
  ) %>% 
  group_by(Tresh,eez_name) %>% 
  summarise(
    ntrans = min(emerging_yr)
  ) %>% 
  group_by(Tresh,ntrans) %>% 
  summarise(
    n_taxa = length(unique(eez_name))
  ) %>% 
  arrange(-n_taxa) %>% 
  View()
  # group_by(Tresh) %>%
  # summarise(sum(n_taxa))
  # ggplot() +
  # geom_line(
  #   aes(
  #     x = ntrans,
  #     y = n_taxa,
  #     color = Tresh
  #   )
  # )

# Numer of EEZs where stocks will emerge
stock_d_hist <- Eme_Res %>% 
  distinct(taxon_key,neighbour_id,Tresh, .keep_all = TRUE) %>%
  group_by(emerging_yr,taxon_key,Tresh) %>%
  summarise(n_tax = length(unique(eez_name))) %>%
  # summarise(n()) %>% 
  mutate(
    level = "EEZs per Spp"
  )  %>% 
  ungroup() %>% 
  select(-taxon_key)

# Per EEZs
eez_d_hist <- Eme_Res %>% 
  distinct(taxon_key,neighbour_id,Tresh, .keep_all = TRUE) %>%
  group_by(emerging_yr,eez_name,Tresh,neighbour_id) %>%  
  # summarise(n_tax = length(unique(taxon_key))) %>% 
  summarise(n()) %>% 
  mutate(
    level = "Spp per EEZ"
  ) %>% 
  ungroup() %>% 
  select(-eez_name)

stock_d_hist %>% 
  bind_rows(eez_d_hist) %>% 
  filter(Tresh != "NA") %>% 
  ggplot() +
  geom_density(
    aes(
      x = emerging_yr,
      fill = level
    ),
    stat = "count",
    alpha= 0.8,
    size=0.5,
  ) +
  scale_y_continuous("Frequency",
                     # breaks = seq(0,300,50),
                     # limits = c(0,300)
  ) +
  scale_x_continuous("Year of emergence",
                     # breaks = seq(2006,2100,25),
                     # limits = c(2000,2100)
  ) +
  my_ggtheme_p() +
  scale_fill_viridis(discrete = T) +
  facet_wrap(~Tresh) +
  ggsave("./Figures/hist_freq.png",
         width = 6,
         height = 6,
         units = "in"
  )

@@

##### Accumulative plots

@@

# Per Stock
stock_d <-  Eme_Res %>% 
  left_join(Neighbours) %>% 
    # distinct(taxon_key,neighbour_id, .keep_all = TRUE) %>% 
    group_by(taxon_key,Tresh) %>% 
    summarise(
      taxon_emer = min(emerging_yr)
    ) %>% 
    group_by(taxon_emer,Tresh) %>% 
  summarise(
    n_tax = n()
  ) %>% 
  group_by(Tresh) %>% 
  mutate(
    com_sum = cumsum(n_tax),
    level = "Stock"
  )

  #  EEZ data
eez_d <-  Eme_Res %>% 
  left_join(Neighbours) %>% 
  filter(n_ens == 10) %>% 
    # distinct(eez_name,neighbour_id, .keep_all = TRUE) %>% 
    group_by(eez_name,Tresh) %>% 
    summarise(
      taxon_emer = min(emerging_yr)
    ) %>% 
    group_by(taxon_emer,Tresh) %>% 
  summarise(
    n_tax = n()
  ) %>% 
  group_by(Tresh) %>% 
  mutate(
    com_sum = cumsum(n_tax),
    level = "EEZ"
  )

# Plot Data
plot_data <- stock_d %>% 
  bind_rows(eez_d) %>% 
  filter(Tresh != "NA") %>% 
  mutate(
    total = ifelse(level == "Stock",length(unique(Eme_Res$taxon_key)),length(unique(Eme_Res$eez_name))),
    per = com_sum/total*100
  ) ;plot_data

  # filter(emerging_yr >= 2020) %>% 
  ggplot() +
    geom_area(data = subset(plot_data, level == "Stock"),
              aes(
                x = as.numeric(taxon_emer),
                y = com_sum,
                fill = level
              ),
              alpha= 0.8,
              size=0.5,
              colour="black"
    )  +
    geom_area(data = subset(plot_data, level == "EEZ"),
              aes(
                x = taxon_emer,
                y = com_sum,
                fill = level
              ),
              alpha= 0.8,
              size=0.5,
              colour="black"
    ) +
    facet_wrap(~Tresh) + 
    # Comment out for option A (Change plot name!!)
  #   geom_hline(yintercept = length(unique(Eme_Res$eez_name)), color = "grey",linetype = "dashed") +
  # geom_text(aes(label="Total number of EEZs", x = 2010, y = length(unique(Eme_Res$eez_name))+10), color = "grey70",size =3) +
  #   geom_hline(yintercept = length(unique(Eme_Res$taxon_key)), color = "grey",linetype = "dashed") +
  # geom_text(aes(label="Total number of Taxa", x = 2010, y = length(unique(Eme_Res$taxon_key))+10), color = "grey70",size =3) +
    scale_y_continuous("Cumulatuve Curve",
                       breaks = seq(0,600,50),
                       limits = c(0,600)
    ) +
    scale_x_continuous("Year of emergence",
                       breaks = seq(2005,2100,10),
                       limits = c(2005,2100)
    ) +
    my_ggtheme_p() +
    scale_fill_viridis(discrete = T) +
  ggsave("./Figures/figure_accumulative_curve_ab.png",
         width = 6,
         height = 6,
         units = "in"
  )

  ### Percentage accumulation

  # filter(emerging_yr >= 2020) %>% 
  ggplot() +
    geom_area(data = subset(plot_data, level == "Stock"),
              aes(
                x = as.numeric(taxon_emer),
                y = per,
                fill = level
              ),
              alpha= 0.8,
              size=0.5,
              colour="black"
    )  +
    geom_area(data = subset(plot_data, level == "EEZ"),
              aes(
                x = taxon_emer,
                y = per,
                fill = level
              ),
              alpha= 0.8,
              size=0.5,
              colour="black"
    ) +
    scale_y_continuous("Cumulatuve Curve",
                       breaks = seq(0,100,25),
                       limits = c(0,100)
    ) +
    scale_x_continuous("Year of emergence",
                       breaks = seq(2005,2100,10),
                       limits = c(2005,2100)
    ) +
    my_ggtheme_p() +
    scale_fill_viridis(discrete = T) +
    facet_wrap(~Tresh) + 
  ggsave("./Figures/figure_accumulative_curve_per.png",
         width = 6,
         height = 6,
         units = "in"
  )

@@

###### Just mean time

```{r emergence_map, eval = T}

# Per EEZs
eez_d_hist <- Eme_Res %>% 
  group_by(eez_name,Tresh) %>% 
  # summarise(n_tax = length(unique(taxon_key))) %>% 
  summarise(
    mean_yr = mean(emerging_yr,na.rm=T),
    sd_yr = sd(emerging_yr,na.rm=T)) %>% 
  gather("metric","value",mean_yr:sd_yr)

### Random plot

ggplot(eez_d_hist) +
  geom_density(
    aes(
      x = value,
      fill = Tresh
    ),
    bins = 10,
    alpha = 0.8
  ) +
  facet_wrap(~metric,
             scales = "free") +
  scale_fill_viridis(discrete = T) +
  my_ggtheme_p() +
  ggsave("./Figures/hist_mean.png",
         width = 6,
         height = 6,
         units = "in"
  )

map <- World_eez_sf %>% 
  left_join(eez_d_hist,
            by = "eez_name") %>% 
  # filter(Tresh == "Tresh_one") %>%
  ggplot() +
  geom_sf(
    aes(
      fill = mean_yr
    )
  ) +
  geom_sf(data =World_map, aes(), size = 0.1, fill = "gray90", col = "gray90") +
  scale_fill_viridis() +
  my_ggtheme_m() +
  facet_wrap(~Tresh,
             ncol = 1) +
  ggsave("./Figures/emer_map_T12.png",
         width = 12,
         height = 10,
         units = "in"
  )

###### With relation to neighbours
